---
title: "HW 3_Group 6_DSCI 5340"
group members: Prasanth Gutha, Manasa Shivkar, Fahimeh Asgari, Vijay Ramaraju Jampana
output:
  pdf_document: default
  html_document: default
---
## Loading the Packages
```{r setup, include=TRUE}
pacman::p_load(caret, data.table, ggplot2, FNN, MASS)
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=6)
theme_set(theme_classic())
options(digits = 3)
```

## 1.Using createDataPartition() function from the Caret package to partition the data into two parts – 80% into training data and 20% into test data.

```{r datapartition}
set.seed(42)
partitionindex <- createDataPartition(y = Boston$medv, p = 0.8, list = FALSE)
train_Boston <- Boston[partitionindex, ]
test_Boston <- Boston[-partitionindex, ]
```
- The above chunk created data partition into two parts. training data set with 407 observation and 14 variables and the test data set with 99 observation and 14 variables.

## 2.Using train() function from the Caret package, run a k-NN model with medv as the response or target variable with
 ## a. Standardize the dataset using center and scale options in the preProcess() function in the Caret package
```{r}
set.seed(42)
prep <- preProcess(train_Boston, method = c("center", "scale"))
train_norm_Boston <- predict(prep, train_Boston)
test_norm_Boston <- predict(prep, test_Boston)
```
 
 ## b. Use a 10-fold cross validation
```{r knn classification}
set.seed(42)
kNN_model <- train(medv ~ .,
                  data = train_norm_Boston,
                  method = "knn",
                  tuneGrid = data.frame(k = 1:18),
                  preProcess = c("center", "scale"),
                  trControl = trainControl(method = "cv", number = 10))
kNN_model
```

## 3.Generate a plot showing model error RMSE against different values of k.
```{r RMSE plot}
ggplot(kNN_model$results, mapping = aes(x = k, y = RMSE)) + 
  geom_point(colour = "black") + 
  geom_line(colour = "purple") +
  ggtitle("Model error RMSE against different values of k")
```

## 4.Choosing the optimal value of k and Explanation
```{r optimal k}
kNN_model$results
kNN_model$bestTune
```

- RMSE was used to select the optimal model using the smallest value. Based on the results from the model, the smallest value is 0.482 when k = 3. We also used best Tune from the model which also suggest the optimal value at k = 3.


